---
title: "Spin-out Data Cleanning - 2017 location"
author: "Meng Ye"
date: "10/10/2023"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

## Introduction 

- update 7/11/2024, update selected columns to include service groups

The current version include the data cleaning for location level CQC data from December 2017. 


### Load packages 

```{r load package, message=FALSE, warning=FALSE}
library(tidyverse) # package for data cleaning and plotting
library(readxl)
library(modelsummary)
library(here) # manage directory
set.seed(5432)
```


## 0. Spin-out list data

```{r message=FALSE, warning=FALSE}
# import CIC spin out coding
spinout <- read_csv(here("data","spinout_list_2017.csv"))
```


## 1. Data at the location level 


```{r import data 0, message=FALSE, warning=FALSE}
# import location level full data
locations<- read_excel(here("data","2017_rating.xlsx"), sheet = 2)
```



### Clean and filter out Gov and CIC providers

```{r rename and filter, message=FALSE, warning=FALSE}
### Rename the columns 

locations_renamed <- locations %>% 
  rename(primary_cat = `Location Primary Inspection Category`,
         type = `Location Type`,
         region = `Location Region`,
         domain = `Key Question`, # named to "Domain" later
         rating = `Latest Rating`,
 #        inherited = `Inherited Rating (Y/N)`, # no such a column in 2017 data
         provider_name = `Provider Name`,
         publication_date = `Publication Date`,
         address = `Location Street Address`) 
```



### Standardize the provider name

```{r sample std}
locations_std <- locations_renamed %>% 
# remove dots
  mutate(std_name = str_replace_all(provider_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC"))
```


### filtering out CIC and GOV providers

```{r}
## building the string detect conditions
keep_terms <- c("CIC", "NHS ", "FOUNDATION TRUST", "COUNCIL", " CC ", " MDC ", " DC ", " BC ", "BOROUGH")
keep_filter <- paste0("\\b(", paste(keep_terms, collapse = "|"), ")\\b")
```

```{r}
## filtering out CICs and GOVs
locations_coded <- locations_std %>% 
  filter(str_detect(std_name, keep_filter)) %>% 
  filter(!str_detect(std_name, "BOROUGH GREEN\\s|BOROUGH ROAD\\s")) %>% 
  ## add the form and level columns
  mutate(form = ifelse(str_detect(std_name, "CIC"), "CIC", "GOV"),
         level = "location") %>% 
  left_join(spinout, by = "provider_name") %>% 
  mutate(spin_out = ifelse(is.na(spin_out), FALSE, TRUE))
nrow(locations_coded)
```


```{r}
#check the spin-out coding distribution 

table(locations_coded$form, locations_coded$spin_out)

```



```{r}
## re-arrange columns 

locations_to_merge <- locations_coded %>%
  mutate(project_id = paste0("location", row_number())) |>
  select(
    project_id,
    std_name,
    provider_name,
    location_name = `Location Name`,
    address,
    type,
    level,
    service_group = `Service / Population Group`,
    domain,
    rating,
    publication_date,
    primary_cat,
    region,
    report_type = `Report Type`,
    form,
    spin_out
  )

```


## 2. Data at the provider level 


```{r import data 0, message=FALSE, warning=FALSE}
# sampled data manually coded (with unfriendly column names)
providers <- read_excel(here("data","2017_rating.xlsx"), sheet = 3)
```


```{r import data, message=FALSE, warning=FALSE}
### Rename the columns 

providers_renamed <- providers %>% 
  rename(primary_cat = `Provider Primary Inspection Category`,
         type = `Provider Type`,
         region = `Provider Region`,
         domain = `Key Question`, # named to "Domain" later 
         rating = `Latest Rating`,
         provider_name = `Provider Name`,
         publication_date = `Publication Date`,
         address = `Provider Street Address`) 
```



### Standardize the provider name

```{r sample std}
providers_std <- providers_renamed %>% 
# remove dots
  mutate(std_name = str_replace_all(provider_name, "\\.", "")) %>% 
# capitalize all
  mutate(std_name = str_to_upper(std_name)) %>%
# standardize to LTD
  mutate(std_name = str_replace_all(std_name, "LIMITED", "LTD")) %>% 
# standardize to CIC
  mutate(std_name = str_replace_all(std_name, "COMMUNITY INTEREST COMPANY", "CIC"))
```


### filtering out CIC and GOV providers

```{r}
## building the string detect conditions
#keep_terms <- c("CIC", "NHS ", "FOUNDATION TRUST", "COUNCIL", " CC ", " MDC ", " DC ", " BC ", "BOROUGH")
#keep_filter <- paste0("\\b(", paste(keep_terms, collapse = "|"), ")\\b")
```

```{r}
## filtering out CICs and GOVs
providers_coded <- providers_std %>% 
  filter(str_detect(std_name, keep_filter)) %>% 
  filter(!str_detect(std_name, "BOROUGH GREEN\\s|BOROUGH ROAD\\s")) %>% 
  ## add the form and level columns
  mutate(form = ifelse(str_detect(std_name, "CIC"), "CIC", "GOV"),
         level = "provider") %>% 
  left_join(spinout, by = "provider_name") %>% 
  mutate(spin_out = ifelse(is.na(spin_out), FALSE, TRUE),
         location_name = NA)
nrow(providers_coded)
```

```{r}
## re-arrange columns 


providers_to_merge <- providers_coded %>%
  mutate(project_id = paste0("provider", row_number())) %>%
  select(
    project_id,
    provider_name,
    location_name,
    address,
    type,
    level,
    service_group = `Service / Population Group`,
    domain,
    rating,
    publication_date,
    primary_cat,
    region,
    report_type = `Report Type`,
    form,
    spin_out
    )
```




## 3. Merge the two level data 

```{r}
combined_2levels <- bind_rows(locations_to_merge, providers_to_merge)
```


## Note: the drop of the overlapping provider level data is conducted in the data analysis file


```{r}
write_rds(combined_2levels, here("data", "spinout_2017.rds"))
```




