---
title: "Spin-out Data Analysis (full data with new grouping) 2 levels 2017 data"
subtitle: "updates to the AOM version"
author: "Meng Ye"
date: "7/3/2024"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

## Notes for edits on 7/3/2024

1. Add the model with health/social care categorization rather than the eight service types
2. add the cross-tab of rating by form and health/social care
3. add random effects model with 8 service types, random effects model with health/social care won't run
4. ***add back*** the dropped provider sheet provider level data

## Notes for editting 1/4/2024
1. I only kept the model we chose to use for ARNOVA 2023
2. The data used in the analysis is the version MERGING in provider sheet data. We keep 2 levels for both government and CICs.
3. Re-ran descriptive statistics to better suit the need for the AOM draft

## Notes for updates 10/17/2023
1. I have checked the false positive CIC "Community Integrated Care (CIC) - 2 Seafarers Walk" found in the 2019 dataset. It is NOT in the 2017 dataset
2. Added data cleaning for dropping overlapping provider rating 
3. Drop certain rating categories as discussed


## Setting up the R session and Loading packages

```{r contr}
# environment setup to run ordered logit properly
options(contrasts = rep("contr.treatment", 2))
```


This chunk loads all the packages to use

```{r load package, message=FALSE, warning=FALSE}
#packages for ordered logit
library(ordinal) # package for ordinal logit regression
library(brant) # brant test for the parallel assumption for ordered logit
library(MASS) # models that work with the brant test
library(tidyverse) # package for data cleaning and plotting
library(readxl) # package for reading excel file
library(broom) # extracting model summary as data frame
library(modelsummary) # deriving model tables
library(scales) # label percent
library(lubridate) # working with dates
library(marginaleffects) #to calculate marginal effects
library(gt) # to format tables
library(here) # work with directory
```

Import the cleaned data 

```{r message=FALSE, warning=FALSE}
merged <- read_rds(here("data", "spinout_2017.rds"))
```


## Prepare the data for ordinal regression

```{r data cleaning}
#select relevant columns, rename and relabel 
merged_cleaned <- merged %>% 
  # recode legal form types to be more readable / easier to present
  mutate(# inherited = ifelse(inherited == "Y", TRUE, FALSE),
         rating = recode(rating, 
                         "Insufficient evidence to rate" = "NA",
                         "Requires improvement" = "Req improv"),
         date = ymd(publication_date)) %>% 
  
  # assign order in the rating levels
  mutate(rating = ordered(rating, levels = c("Inadequate","Req improv", "Good", "Outstanding")),
         form_spinout = case_when(form == "GOV" ~ "GOV",
                           form == "CIC" & spin_out == "TRUE" ~ "SP_CIC",
                           form == "CIC" & spin_out == "FALSE" ~ "Ind_CIC"),
         form_spinout = fct_relevel(form_spinout, "GOV"),
         social_care = ifelse(type == "Social Care Org", "social care", "healthcare")) %>% 
  
  
  # creating a new dummy variable for facility category
  mutate(year = year(date),
         year2 = year-2013,
         Year = factor(year)) %>%
  
  # regroup care type

  mutate(service_type = case_when(
      str_detect(primary_cat, "Acute hospital") ~ "Acute hospital",
      str_detect(primary_cat, "Mental health") ~ "Mental health",
      TRUE ~ primary_cat  # Keep the original value if none of the conditions match
    )) %>%
  
  # drop unmatched service types
  filter(!service_type %in% c("Acute hospital", "Ambulance service", "Dentists", 
                              "Independent consulting doctors", "Prison Healthcare")) %>% 
  
  # drop overlapping provider data 
  # commented out this section because decided to keep provider and coreservice ratings in the provider sheet to stay consistant with the location sheet data
#  group_by(provider_name) %>% 
#  mutate(overlap = ifelse(
#    n() > 1 & level == "provider" & report_type == "Provider", "drop", "keep"
#  )) %>%
#  ungroup()  %>% 
#  filter(overlap != "drop") %>% 
  # 11784 rolws of data down to 11310 rows
  
  # converting the ordinal variable to numerical 
  mutate(rating_num = case_when(rating == "Inadequate" ~ 1,
                                rating == "Req improv" ~ 2,
                                rating == "Good" ~ 3,
                                rating == "Outstanding" ~ 4)) 
nrow(merged_cleaned)
```

```{r}
merged_cleaned %>% 
  mutate(overall = ifelse(domain == "Overall", 1, 0)) %>% 
  group_by(form_spinout) %>% 
  summarize(count_provider = n_distinct(provider_name),
            count_location = n_distinct(location_name),
            count_overall_rating = sum(overall),
            count_rating = n()) %>% 
  gt()
```

**Note** The count of unique location name does not include provider level data lines. 



## Dverall distribution of key variables - overall

```{r}
datasummary(form_spinout + social_care + service_type  + region + Year ~ 1, data = merged_cleaned, fmt = 0)
```


## Distribution of key variables - three groups 


```{r}
datasummary(1 + rating + social_care + service_type + region ~ form_spinout, data = merged_cleaned, fmt = 0)
```

## Cross-tab of rating data 

```{r}
datasummary_crosstab(form_spinout * social_care ~ rating, data = merged_cleaned)
```
## Calculating meaning rating score 

```{r}
merged_overall <- merged_cleaned %>% 
  filter(domain == "Overall")
datasummary(form_spinout * social_care ~ rating_num * mean, data = merged_overall)
```
```{r}
datasummary(service_type ~ form_spinout * rating_num * mean, data = merged_overall)

```


## whole models with continous year trend  - fixed effects

Since service_type is inclusive of variation/information from `socialcare` and `level`, we drop this variable from the model 


```{r }
model_order_overall <- clm(rating ~ form_spinout  + service_type + region + year2,
                data = filter(merged_cleaned, domain == "Overall"),
                link = "logit")

model_order_safe <- clm(rating ~ form_spinout  + service_type + region + year2,
                data = filter(merged_cleaned, domain == "Safe"),
                link = "logit")
model_order_effective <- clm(rating ~ form_spinout  + service_type + region + year2,
                data = filter(merged_cleaned, domain == "Effective"),
                link = "logit")
model_order_caring <- clm(rating ~ form_spinout  + service_type + region + year2,
                data = filter(merged_cleaned, domain == "Caring"),
                link = "logit")
model_order_well_led <- clm(rating ~ form_spinout + service_type + region + year2,
                data = filter(merged_cleaned, domain == "Well-led"),
                link = "logit")
model_order_responsive <- clm(rating ~ form_spinout + service_type + region + year2,
                data = filter(merged_cleaned, domain == "Responsive"),
                link = "logit")
```

### Compiling all results in odds ratio (exponentiated)

```{r}
ordinal_models_exp <-
  modelsummary(
    list(
      "overall" = model_order_overall,
      "safe" = model_order_safe,
      "effective" = model_order_effective,
      "caring" = model_order_caring,
      "well-led" = model_order_well_led,
      "responsive" = model_order_responsive
    ),
    coef_omit = "region",
    exponentiate = T,
    statistic = "({p.value}) {stars}")
ordinal_models_exp
```

## whole models with continous year trend  - random effects (multi-level)


```{r }
model_order_overall_multi <-
  clmm(
    rating ~ form_spinout  + (1 | service_type) + (1 | region) + year2,
    data = filter(merged_cleaned, domain == "Overall"),
    link = "logit"
  )

model_order_safe_multi <-
  clmm(
    rating ~ form_spinout  + (1 | service_type) + (1 | region) + year2,
    data = filter(merged_cleaned, domain == "Safe"),
    link = "logit"
  )
model_order_effective_multi <-
  clmm(
    rating ~ form_spinout  + (1 | service_type) + (1 | region) + year2,
    data = filter(merged_cleaned, domain == "Effective"),
    link = "logit"
  )
model_order_caring_multi <-
  clmm(
    rating ~ form_spinout  + (1 | service_type) + (1 | region) + year2,
    data = filter(merged_cleaned, domain == "Caring"),
    link = "logit"
  )
model_order_well_led_multi <-
  clmm(
    rating ~ form_spinout + (1 | service_type) + (1 | region) + year2,
    data = filter(merged_cleaned, domain == "Well-led"),
    link = "logit"
  )
model_order_responsive_multi <-
  clmm(
    rating ~ form_spinout + (1 | service_type) + (1 | region) + year2,
    data = filter(merged_cleaned, domain == "Responsive"),
    link = "logit"
  )
```


### Compiling all results in odds ratio (exponentiated)


```{r}
ordinal_models_exp_multi <-
  modelsummary(
    list(
      "overall" = model_order_overall_multi,
      "safe" = model_order_safe_multi,
      "effective" = model_order_effective_multi,
      "caring" = model_order_caring_multi,
      "well-led" = model_order_well_led_multi,
      "responsive" = model_order_responsive_multi
    ),
    coef_omit = "region",
    exponentiate = T,
    statistic = "({p.value}) {stars}")
ordinal_models_exp_multi
```

## whole models with continous year trend by social/health care

Since service_type is inclusive of variation/information from `socialcare` and `level`, we drop this variable from the model 


### Compiling all results in odds ratio (exponentiated)



```{r }
model_order_overall2 <- clm(rating ~ form_spinout + social_care + region + year2,
                data = filter(merged_cleaned, domain == "Overall"),
                link = "logit")

model_order_safe2 <- clm(rating ~ form_spinout  + social_care + region + year2,
                data = filter(merged_cleaned, domain == "Safe"),
                link = "logit")
model_order_effective2 <- clm(rating ~ form_spinout  + social_care + region + year2,
                data = filter(merged_cleaned, domain == "Effective"),
                link = "logit")
model_order_caring2 <- clm(rating ~ form_spinout  + social_care + region + year2,
                data = filter(merged_cleaned, domain == "Caring"),
                link = "logit")
model_order_well_led2 <- clm(rating ~ form_spinout + social_care + region + year2,
                data = filter(merged_cleaned, domain == "Well-led"),
                link = "logit")
model_order_responsive2 <- clm(rating ~ form_spinout + social_care + region + year2,
                data = filter(merged_cleaned, domain == "Responsive"),
                link = "logit")
```



```{r}
ordinal_models_exp2 <-
  modelsummary(
    list(
      "overall" = model_order_overall2,
      "safe" = model_order_safe2,
      "effective" = model_order_effective2,
      "caring" = model_order_caring2,
      "well-led" = model_order_well_led2,
      "responsive" = model_order_responsive2
    ),
    coef_omit = "region",
    exponentiate = T,
    statistic = "({p.value}) {stars}")
ordinal_models_exp2
```


